---
title: "Day 7 Homework"
output: html_document
---

Find three relationships in your data.  For each relationship, do the following:

1. Verbally articulate the relationship you want to examine with a plot.
# Is teen self reported physiological arousal related to teen self reported anxiety as measured by the SPAI-C? 
# Do teens who belong in the clinical sample report having significantly higher arousal than teens in the community sample? 
# We can also group teens by classifying them along cut off scores (at or above 18) for the SPAI-C, suggesting they have probable social anxiety disorder. I will make a new grouping variable to sort each participant. Are teens who scored above the cut off scores scoring higher on the SAFE than teens who were below the cut off score? 

2. Verbally identify what variables you want to map to each graphical dimension.
# X axis: teen SPAI-C score; Y axis; teen arousal 
# X axis: group status; Y axis: teen arousal
# X axis: SPAI-C cut off group; Y axis: teen SAFE scores

3. Reshape your data as needed.

```{r}
#libraries
library(haven)
library(ggplot2)
library(tidyr)
library(dplyr)

#reading dataset
SAS.data <- read_sav("QASMIEH_DATASET_PSYC798W.sav")

#making dataframe of only variables of interest
my.SAS <- c( "MASTER_ID","GROUPAB" ,"SAFE_TOTAL_C","spaicctot","TEEN_SAM_COMP_ALL_TASKS")

SAS.df <- SAS.data[,my.SAS]
SAS.df$GROUPAB <- as.factor(SAS.df$GROUPAB)
levels(SAS.df$GROUPAB) <- c("Community sample", "Clinical sample")
levels(SAS.df$GROUPAB)


#making new variable
SAS.df$SPAICCUTOFF <- NA
SAS.df$SPAICCUTOFF[SAS.df$spaicctot<18] <- "Below cutoff"
SAS.df$SPAICCUTOFF[SAS.df$spaicctot>=18] <- "Above cutoff"
SAS.df$SPAICCUTOFF<- as.factor(SAS.df$SPAICCUTOFF)
summary(SAS.df)

#reordering so factors variables are first
SAS.df <- SAS.df[c(1,2,6,3,4,5)]

#creating long dataset
SAS.long <- gather (SAS.df, "Measure", "Response", SAFE_TOTAL_C:TEEN_SAM_COMP_ALL_TASKS)
SAS.long

```

4. Use ggplot2 to create your graphs.

```{r}
# Is teen self reported physiological arousal related to teen self reported anxiety as measured by the SPAI-C? 
(arousalanxiety <- ggplot(SAS.df, aes(TEEN_SAM_COMP_ALL_TASKS, spaicctot))+geom_point(aes(color=GROUPAB)))

# Do teens who belong in the clinical sample report having significantly higher arousal than teens in the community sample? 
mean.arousal.A <-mean(SAS.df$TEEN_SAM_COMP_ALL_TASKS[SAS.df$GROUPAB=="Community sample"])
mean.arousal.B <-mean(SAS.df$TEEN_SAM_COMP_ALL_TASKS[SAS.df$GROUPAB=="Clinical sample"])
grouping <- c("Community sample", "Clinical sample")
means.arousal <- c(mean.arousal.A, mean.arousal.B)
grouparousal <- data.frame(grouping, means.arousal)

(arousalplot <- ggplot(grouparousal, aes(grouping, means.arousal))+geom_bar(stat="identity", aes(fill=grouping)))

# We can also group teens by classifying them along cut off scores (at or above 18) for the SPAI-C, suggesting they have probable social anxiety disorder. I will make a new grouping variable to sort each participant. Are teens who scored above the cut off scores scoring higher on the SAFE than teens who were below the cut off score? 
(mean.SAFE.Above <-mean(SAS.df$SAFE_TOTAL_C[SAS.df$SPAICCUTOFF=="Above cutoff"]))
(mean.SAFE.Below <-mean(SAS.df$SAFE_TOTAL_C[SAS.df$SPAICCUTOFF=="Below cutoff"]))
groupingSAFE <- c("Above cutoff", "Below cutoff")
means.SAFE <- c(mean.SAFE.Above, mean.SAFE.Below)
SAFEcut <- data.frame(groupingSAFE, means.SAFE)

(SAFEplot <- ggplot(SAFEcut, aes(groupingSAFE, means.SAFE))+geom_bar(stat="identity", aes(fill=groupingSAFE)))
```

5. Fiddle with colors, axis labels, themes, etc. until the plot looks "publication ready" to you.  Try to change at least one thing per plot.

```{r}
new1<- arousalanxiety+
       xlab("Teen Self Report Arousal")+
       ylab("Teen Self Reported Anxiety (SPAI-C)")+
       ggtitle("Teen Arousal and Anxiety")+
       scale_color_brewer(palette = "Pastel1", name="Sample Grouping")+
       theme(plot.title = element_text(hjust = 0.5))

new2<- arousalplot+
       xlab("Sample grouping")+
       ylab("Mean Teen Self Report Arousal")+
       ggtitle("Teen Arousal by Sample Group")+
       scale_fill_brewer(palette = "Spectral", name="Sample Grouping")+
       theme(plot.title = element_text(hjust = 0.5)) 

new3<- SAFEplot+
       xlab("SPAI-C Cutoff Group")+
       ylab("Mean Teen Self Reported Safety Behaviors (SAFE)")+
       ggtitle("Teen Safety Behaviors by SPAI-C Cutoff Groupings")+
       scale_fill_brewer(palette = "Dark2",name="SPAI-C Cutoff Grouping")+
       theme(plot.title = element_text(hjust = 0.5))

```

6. Save all plots to a format that you could use in a paper.  HINT: PNG works well for pasting into Word documents.

```{r}
png("HW7plot1.png")
plot(new1)
dev.off()
png("HW7plot2.png")
plot(new2)
dev.off()
png("HW7plot3.png")
plot(new3)
dev.off()
png("HW7plot4.png")
plot(arousalanxiety)
dev.off()
png("HW7plot5.png")
plot(arousalplot)
dev.off()
png("HW7plot6.png")
plot(SAFEplot)
dev.off()
```

